Kids PM Summary

### 2019-12-20 dustrak_data_cleaning.Rmd
PM cleaning

---
title: "KidsAIR initial PM data work"
author: "Ethan Walker"
date: "Started 19 Nov 2019, Updated 19 Dec 2019"
output: pdf_document
---

@@

# Initial load, format, and save KidsAIR Dusttrak data
```{r}
##### Read in dusttrack .txt files (from WMT) #####

@@

# Change function input to look at specific home
# Run all of these lines together
pm_time_plot_function(id = "137", location = "NN")
head(gaps, 25)
head(pm_desc, 25)
tail(pm_desc, 25)

# filter data for a single home
check_data <- kids_pm %>% 
  filter(area == "NN" & home_winter_id == "140")

%>% 
  filter(pm < 0)
```

@@

kids_pm_new2 <- kids_pm_new %>% 

@@

  distinct(time_diff, .keep_all = TRUE)

kids_pm_new2 <- kids_pm_new %>%
  ungroup() %>% 
  arrange(area, home_winter_id, pm_datetime_new)

# write_rds(kids_pm_new2, "Output/kids_pm_new.rds")
```

@@

# Original code/files from Emily - save for future reference
```{r}
##### Load combined dusttrack files from Emily W #####
##### Compare these with the raw data pulled in above #####

## List sheets within the excel file
list_sheets_nn <- excel_sheets("Input/nn_kids_dusttrack_20191121.xlsx")
## Apply the read_xlsx function to each item in the sheet list
## Bind the items from each sheet into a single dataframe
pm_initial_nn <- data.frame(do.call("rbind", lapply(list_sheets_nn, read_xlsx, 
                                      path = "Input/nn_kids_dusttrack_20191121.xlsx"))) %>% 
  mutate(area = "NN")

list_sheets_wmt <- excel_sheets("Input/wmt_kids_dusttrack_20191121.xlsx")
pm_initial_wmt <- data.frame(do.call("rbind", lapply(list_sheets_wmt, read_xlsx, 
                                      path = "Input/wmt_kids_dusttrack_20191121.xlsx"))) %>% 
  mutate(area = "WMT")

# Bind data together from different study areas
pm_initial <- rbind(pm_initial_nm, pm_initial_wmt)

# Rename and create new variables
kids_pm <- pm_initial %>% 
  # separate time from incorrect/associated date
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  # format/rename variables
  mutate(home_winter_id = as.factor(homeid),
         pm_date = ymd(date),
         pm_time = hms(time),
         pm = as.numeric(aerosol)) %>% 
  # new datetime variable
  unite(pm_datetime, c("date", "time"), sep = " ", remove = FALSE) %>% 
  select(home_winter_id, area, pm_date, pm_time, pm_datetime, pm) %>% 
  arrange(home_winter_id, pm_date) %>% 
  group_by(home_winter_id) %>% 
  # new variables summarizing pm2.5 within each home_winter_id
  mutate(pm_total_observations = n(),
         pm_total_summed = sum(pm),
         min_datetime = first(pm_datetime),
         max_datetime = last(pm_datetime),
         min_datetime = ymd_hms(min_datetime),
         max_datetime = ymd_hms(max_datetime),
         diff_datetime = interval(min_datetime, max_datetime),
         pm_total_days = diff_datetime/86400,
         pm_24hr_mean = pm_total_summed/pm_total_days,
         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
         obs = row_number(),
         sampling_day = if_else(obs < 1441, 1, 0),
         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
  group_by(home_winter_id, sampling_day) %>% 
  mutate(pm_24hr_daily = sum(pm),
         pm_24hr_daily_weighted = pm_24hr_daily*(1440/n()),
         pm_sampling_day = sampling_day) %>% 
  ungroup() %>% 
  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day)

# Save pm dataset as RDS and CSV
write_rds(kids_pm, "Output/kids_pm.rds")
write_csv(kids_pm, "Output/kids_pm.csv")
```

>>>>

---
title: "KidsAIR initial PM data work"
author: "Ethan Walker"
date: "Started 19 Nov 2019, Updated 20 Dec 2019"
output: pdf_document
---

@@

# Initial load, format, and save KidsAIR Dusttrak data
## Files here are in 3 formats: csv (2 different formats) and txt
## I had to do some manual cleaning of the files to get them named
## consistently and in reasonably similar formats to be able to load
## into R in batches, as seen below. The files I upload in the code below
## will also be uploaded to Box in a folder with my name.
## Box folder directory: 
```{r}
##### Read in dusttrack .txt files (from WMT) #####

@@

# Change function input to look at specific home
# Run all of these lines together
pm_time_plot_function(id = "129", location = "WMT")
head(gaps, 25)
head(pm_desc, 25)
tail(pm_desc, 25)

# filter data for a single home
check_data <- kids_pm %>% 
  filter(area == "WMT" & home_winter_id == "129") %>% 
  filter(pm < 0)
```

@@

kids_pm_new <- kids_pm %>% 

@@

  ungroup() %>% 
  arrange(area, home_winter_id, pm_datetime_new)

@@

### 2019-12-20 pm_summary.Rmd
PM cleaning

# Summary stats
```{r}
pm_summary <- kids_pm %>% 
  # group by: area, treatment, dtid, day_of_week, Sampling_day
  group_by(day_of_week) %>% 
  summarize("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE),
            "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
            "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE),
            "Mean % Daily Obs" = mean(percent_daily_obs_pm, na.rm = TRUE))
```

# Boxplots for PM by study area
```{r, fig.height=6, fig.width=5}
boxplots_kids_pm <- kids_pm %>% 
  #filter(pm < 500) %>% 
  ggplot(aes(area, log(pm))) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    #ylim(0, 500) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_pm
```

>>>>

# Summary stats
```{r}
pm_summary <- kids_pm %>% 
  # group by: area, treatment, dtid, day_of_week, Sampling_day
  group_by(sampling_day) %>% 
  summarize("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE),
            "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
            "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE),
            "Mean % Daily Obs" = mean(percent_daily_obs_pm, na.rm = TRUE))
```

# Boxplots for PM by study area
```{r, fig.height=6, fig.width=5}
boxplots_kids_pm <- kids_pm %>% 
  #filter(pm < 500) %>% 
  ggplot(aes(area, pm)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    ylim(0, 100) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_pm
```
